---
title: "Vignette Title"
author: "Johannes Signer, Femke Pflüger, Niko Balkenhol"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Intro

Developements of the `fpga` packages were heavily inspired by `gstudio`. While `gstudio` provides all of the functions and much more use friendliness, we were faced with the challenge to analyse several thousands of indivudals from hunderts of populations and `gstudio` was to slow. This vignette severs as a comparison between `fpga` and `gstudio`.

# Data

We use the `arapat` data from `gstudio`:

```{r}
library(abind)
library(fpga)
library(gstudio)


# We have 
# - 1000 Individuals
# - with 3 loci
# - 200 possible allels
# - in 10 populations

set.seed(111)
dat0 <- data.frame(
  pop = factor(as.character(rep(1:9, each = 100))), 
  loc1 = do.call(c, lapply(split(sample(c(1:100), 900 * 2, TRUE), rep(1:900, each = 2)), locus)), 
  loc2 = do.call(c, lapply(split(sample(c(200:250), 900 * 2, TRUE), rep(1:900, each = 2)), locus)), 
  loc3 = do.call(c, lapply(split(sample(c(1:200), 900 * 2, TRUE), rep(1:900, each = 2)), locus))
)


# data for fpga
x <- lapply(column_class(dat0, class = "locus"), function(i) apply(alleles(dat0[[i]]), 2, as.integer))
dat1 <- abind(list(sapply(x, function(y) y[, 1]), 
                  sapply(x, function(y) y[, 2])), along = 3)
```

# Genetic Metrics

Number of alleles: 

```{r}
system.time(na0 <- A(dat0))
system.time(na1 <- n_alls(dat1))

na0$A == na1
```

Observed heterozygosity:

```{r}
system.time(ho0 <- Ho(dat0))
system.time(ho1 <- het_obs(dat1))
ho0$Ho == ho1
```

Expected heterozygosity:

```{r}
system.time(he0 <- He(dat0))
system.time(he1 <- het_exp(dat1))
he0$He == he1
```


Effective number of allels:


```{r}
system.time(na0 <- Ae(dat0))
system.time(na1 <- n_eff_alls(dat1))

na0$Ae == na1
```


## Adding a stratum

All of the above function can be stratified by population, e.g., 

```{r}
system.time(na0 <- genetic_diversity(dat0, stratum = "pop", mode = "A"))
system.time(na1 <- n_alls(dat1, dat0[, 1]))
sum(na0$A) == sum(na1)
```


# Genetic structure

## `gst`

```{r}
system.time(gst0 <- genetic_structure(dat0, stratum = "pop", mode = "Gst", pairwise = FALSE, size.correct = FALSE))  ## 0.35 sec
system.time(gst1 <- gst(dat1, stratum = dat0$pop, size.correct = FALSE))  ## 0.35 sec
all(gst0$Gst == gst1[1, ])

system.time(gst0 <- genetic_structure(dat0, stratum = "pop", mode = "Gst", pairwise = FALSE, size.correct = TRUE))  ## 0.35 sec
system.time(gst1 <- gst(dat1, stratum = dat0$pop, size.correct = TRUE))  ## 0.35 sec
all(gst0$Gst == gst1[1, ])


system.time(gst0 <- genetic_structure(dat0, stratum = "pop", mode = "Gst", pairwise = TRUE, size.correct = FALSE))  ## 0.35 sec
system.time(gst1 <- pw_gstructure(dat1, stratum = dat0$pop, size.correct = FALSE, what = "gst"))  ## 0.35 sec
sum(gst0, na.rm = TRUE) == sum(gst1, na.rm = TRUE)

system.time(gst0 <- genetic_structure(dat0, stratum = "pop", mode = "Gst", pairwise = TRUE, size.correct = TRUE))  ## 0.35 sec
system.time(gst1 <- pw_gstructure(dat1, stratum = dat0$pop, size.correct = TRUE, what = "gst"))  ## 0.35 sec
sum(gst0, na.rm = TRUE) == sum(gst1, na.rm = TRUE)
```

## `dest`

```{r}
system.time(gst0 <- genetic_structure(dat0, stratum = "pop", mode = "Dest", pairwise = FALSE, size.correct = FALSE))  ## 0.35 sec
system.time(gst1 <- dest(dat1, stratum = dat0$pop, size.correct = FALSE))  ## 0.35 sec
all(gst0$Dest == gst1[1, ])

system.time(gst0 <- genetic_structure(dat0, stratum = "pop", mode = "Dest", pairwise = FALSE, size.correct = TRUE))  ## 0.35 sec
system.time(gst1 <- dest(dat1, stratum = dat0$pop, size.correct = TRUE))  ## 0.35 sec
all(gst0$Dest == gst1[1, ])


system.time(gst0 <- genetic_structure(dat0, stratum = "pop", mode = "Gst", pairwise = TRUE, size.correct = FALSE))  ## 0.35 sec
system.time(gst1 <- pw_gstructure(dat1, stratum = dat0$pop, size.correct = FALSE, what = "gst"))  ## 0.35 sec
sum(gst0, na.rm = TRUE) == sum(gst1, na.rm = TRUE)

system.time(gst0 <- genetic_structure(dat0, stratum = "pop", mode = "Gst", pairwise = TRUE, size.correct = TRUE))  ## 0.35 sec
system.time(gst1 <- pw_gstructure(dat1, stratum = dat0$pop, size.correct = TRUE, what = "gst"))  ## 0.35 sec
sum(gst0, na.rm = TRUE) == sum(gst1, na.rm = TRUE)
```


# Pairwise genetic structure

## GST
```{r}

system.time(gst0 <- genetic_structure(dat0, stratum = "pop", mode = "Gst", pairwise = TRUE, size.correct = TRUE))  ## 0.35 sec
system.time(gst1 <- pw_gstructure(dat1, stratum = dat0$pop, what = "gst", size.correct = TRUE))  ## 0.35 sec

sum(gst0, na.rm = TRUE)
sum(gst1, na.rm = TRUE)


system.time(gst0 <- genetic_structure(dat0, stratum = "pop", mode = "Gst", pairwise = TRUE, size.correct = TRUE))  ## 0.35 sec
system.time(gst1 <- pw_gstructure(dat1, stratum = dat0$pop, size.correct = TRUE, what = "gst"))  ## 0.35 sec

```

