---
title: "Vignette Title"
author: "Johannes Signer, Femke Pflüger, Niko Balkenhol"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



# Intro

Developements of the `fpga` packages were heavily inspired by `gstudio`. While `gstudio` provides all of the functions and much more use friendliness, we were faced with the challenge to analyse several thousands of indivudals from hunderts of populations and `gstudio` was to slow. This vignette severs as a comparison between `fpga` and `gstudio`.

# Data

We use the `arapat` data from `gstudio`:



```{r}
library(abind)
library(fpga)
library(gstudio)


# We have 
# - 1000 Individuals
# - with 3 loci
# - 200 possible allels
# - in 10 populations

set.seed(111)
dat0 <- data.frame(
  pop = factor(as.character(rep(1:9, each = 100))), 
  loc1 = do.call(c, lapply(split(sample(c(1:100, NA), 900 * 2, TRUE), rep(1:900, each = 2)), locus)), 
  loc2 = do.call(c, lapply(split(sample(c(200:250, NA), 900 * 2, TRUE), rep(1:900, each = 2)), locus)), 
  loc3 = do.call(c, lapply(split(sample(c(1:200), 900 * 2, TRUE), rep(1:900, each = 2)), locus))
)

## With NA
dat0 <- data.frame(
  pop = factor(as.character(rep(1:3, each = 10))), 
  loc1 = do.call(c, lapply(split(sample(c(1:10, NA), 30 * 2, TRUE), rep(1:30, each = 2)), locus)), 
#  loc2 = do.call(c, lapply(split(sample(c(200:205, NA), 30 * 2, TRUE), rep(1:30, each = 2)), locus)), 
  loc3 = do.call(c, lapply(split(sample(c(1:4), 30 * 2, TRUE), rep(1:30, each = 2)), locus))
)

# No NA
dat0 <- data.frame(
  pop = factor(as.character(rep(1:3, each = 10))), 
  loc1 = do.call(c, lapply(split(sample(c(1:10), 30 * 2, TRUE), rep(1:30, each = 2)), locus)), 
#  loc2 = do.call(c, lapply(split(sample(c(200:205, NA), 30 * 2, TRUE), rep(1:30, each = 2)), locus)), 
  loc3 = do.call(c, lapply(split(sample(c(1:4), 30 * 2, TRUE), rep(1:30, each = 2)), locus))
)

head(dat0)

dat2 <- gstudio2fpga(dat0)
identical(dat1, dat2)
dat1
dat2
dat00 <- fpga2gstudio(dat1)
identical(dat0, dat00)


# data for fpga
x <- lapply(column_class(dat0, class = "locus"), function(i) apply(alleles(dat0[[i]]), 2, as.integer))
dat1 <- abind(list(sapply(x, function(y) y[, 1]), 
                  sapply(x, function(y) y[, 2])), along = 3)
```

# Genetic Metrics

Frequencies

```{r}
gstudio::frequencies(dat0)
freq(dat1)

mean(frequencies(dat0)$Frequency) == mean(freq(dat1)$freq)
sum(frequencies(dat0)$Frequency) 
sum(freq(dat1)$freq)


```


Number of alleles: 

```{r}
system.time(na0 <- A(dat0))
system.time(na1 <- n_alls(dat1))

na0$A == na1
```

Observed heterozygosity:

```{r}
system.time(ho0 <- Ho(dat0))
system.time(ho1 <- het_obs(dat1))
ho0$Ho == ho1
```

Expected heterozygosity:

```{r}
system.time(he0 <- He(dat0))
system.time(he1 <- het_exp(dat1))
he0$He == he1
```


Effective number of allels:


```{r}
system.time(na0 <- Ae(dat0))
system.time(na1 <- n_eff_alls(dat1))

na0$Ae == na1
```


## Adding a stratum

All of the above function can be stratified by population, e.g., 

```{r}
system.time(na0 <- genetic_diversity(dat0, stratum = "pop", mode = "A"))
system.time(na1 <- n_alls(dat1, dat0[, 1]))
sum(na0$A) == sum(na1)
```


# Genetic structure

## `gst`


## `dest`



# Pairwise genetic structure

## GST
