---
title: "Introduciton to `fpga`"
author: "Johannes Signer, Femke Pflüger, Niko Balkenhol"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


# What is the package

- calculation of pop genetic measures

# Which data are needed

- diploid only

# Comparison to other packages

- optimized for speed at the cost of usabilitiy

- adegenet and gstudio: much more sophisticated representation of genetic data (we only use 3d arrays)
- both packages offer many more methods than we do
- because of their useability, there is some overhead interms of data checking, which we try to avoid here
- PopGenReport most user friendly
- 

# Example 

Make sure to get the newest version of the package:

```{r, eval = FALSE}
devtools::install_github("jmsigner/fpga")
```


## Reading data

### From scratch

It is relatively straight forward to create suitable datat structures from a `data.frame` using the `fpga::df2fpga` function.

```{r}
# fpga::df2fpga # does not exist yet 

```


### From files 

Currently there only `fstat` is support:

```{r}
dat1 <- fpga::read_fstat(system.file('sample-data/simulation_s1e2.dat', package='fpga'))
```


### From other packages

`genind` objects from adegenet can be converted using:


```{r, eval = FALSE}
dat <- adegenet::read.fstat(system.file('sample-data/simulation_s1e2.dat', package='fpga'))
dat2 <- fpga::genind2fpga(dat) # some more checks needed (some alleles are are switched)
```

similarely from `gstudio` 

```{r, eval = FALSE}
fpga::gstudio2fpga
```

# Population genetic measures (simulated data, no missing data)

We use a simulated data set:

```{r}
library(fpga)
dat0 <- fpga::read_fstat(system.file('sample-data/simulation_s1e2.dat', package='fpga'))
dat1 <- adegenet::read.fstat(system.file('sample-data/simulation_s1e2.dat', package='fpga'))
```


## Number of allels

```{r}
n_alls(dat0, rownames(dat0))
n_eff_alls(dat0, rownames(dat0))
```

## Number of private allels

```{r}
fpga::nprivate_alleles(dat0, rownames(dat0))
```

## Observed and expected heterzygosity

```{r}
fpga::het_exp(dat0)
fpga::het_obs(dat0)
```

In adegenet

```{r}
summary(dat1) 
```

By population

```{r}
fpga::het_exp(dat0, rownames(dat0))
fpga::het_obs(dat0, rownames(dat0))
```


## $G_{ST}$

```{r}

fpga::gst(dat0, rownames(dat0))
mmod::Gst_Nei(dat1)
```

## $G_{ST}'$

```{r}
fpga::gst_p(dat0, rownames(dat0))
mmod::Gst_Hedrick(dat1)
```
## $D_{ST}'$

```{r}
fpga::dest(dat0, rownames(dat0))
mmod::D_Jost(dat1)
```

# Genetic differentiation (real data)

We use the `nancycat` data set from the adegenet package for illustration. 

```{r}
library(fpga)
data("nancycats", package = "adegenet")
dat <- genind2fpga(nancycats)
str(dat)
```

## Population genetic measures

We use the `nancycat` data set from the adegenet package for illustration. 

```{r}
library(fpga)
data("nancycats", package = "adegenet")
dat <- genind2fpga(nancycats)
str(dat)
```


## Number of allels

```{r}
n_alls(dat, rownames(dat))
n_eff_alls(dat, rownames(dat))
```

## Number of private allels

```{r}
fpga::nprivate_alleles(dat, rownames(dat))
```

## Observed and expected heterzygosity

```{r}
fpga::het_exp(dat)
fpga::het_obs(dat)
```

In adegenet

```{r}
adesum <- summary(nancycats) ## results differ at time, most likely due to missing values
summary(nancycats) ## results differ at time, most likely due to missing values
```

By population

```{r}
fpga::het_exp(dat, rownames(dat))
fpga::het_obs(dat, rownames(dat))
```


# Genetic differentiation

## $G_{ST}$

```{r}
fpga::gst(dat, rownames(dat))
mmod::Gst_Nei(nancycats)
```

## $G_{ST}'$

```{r}
fpga::gst_p(dat, rownames(dat))
mmod::Gst_Hedrick(nancycats)
```
## $D_{ST}'$

```{r}
fpga::dest(dat, rownames(dat))
mmod::D_Jost(nancycats)
```


